Concurrent Programming in Harmony

### 2020-09-30 book/paper.tex
Latest

Tony Hoare, who came up with the concept of split binary semaphores, devised
an abstraction of the concept in a programming language paradigm called
\emph{monitors}~\cite{Hoare74}.
\index{monitor}%
(A similar construct was independently invented by Per Brinch Hansen~\cite{BH73}.)
A monitor is a special version of an object-oriented \emph{class}, comprising
a set of variables and methods that operate on those variables.
There is a split binary semaphore associated with each such class.
The \textit{mutex} is hidden: it is automatically acquired when invoking a
method and released upon exit.
The other semaphores are encapsulated in \emph{condition variables}.
\index{condition variable}%
Each condition variable also keeps track of the number of processes waiting
on the condition variable.
There are two operations on condition variables: \texttt{wait}
\index{wait}%
and
\texttt{signal}.
\index{signal}%

\autoref{fig:hoare} presents the semantics by giving an implementation of
\texttt{wait} and \texttt{signal} as well as the code that is used to enter
and exit monitor methods.
\autoref{fig:hoaretest} presents a simple example lifted directly from
the original paper~\cite{Hoare74}.
In the design proposed by Hoare, the monitor methods automatically acquire
and release the monitor mutex.

\texttt{wait} increments the condition's counter, releases the monitor mutex,
blocks while trying to procure the condition's semaphore, and finally decrements
the counter---in much the same way as we have seen for SBS.
\texttt{signal} checks to see if the condition's count is non-zero, and if so
vacates the condition's semaphore---passing the baton to
a process trying to procure it---and then blocks by procuring the mutex again.
The use of \texttt{wait} and \texttt{signal} ensures that \texttt{P} and
\texttt{V} operations alternate as prescribed by the SBS technique.
Our implementations of reader/writer locks can be easily changed to
use \texttt{wait} and \texttt{signal} instead of using \texttt{P}
and \texttt{V}.

>>>>

Tony Hoare, who came up with the concept of split binary semaphores, devised
an abstraction of the concept in a programming language paradigm called
\emph{monitors}~\cite{Hoare74}.
\index{monitor}%
(A similar construct was independently invented by Per Brinch Hansen~\cite{BH73}.)
A monitor is a special version of an object-oriented \emph{class}, comprising
a set of variables and methods that operate on those variables.
There is a split binary semaphore associated with each such class.
The \textit{mutex} is hidden: it is automatically acquired when invoking a
method and released upon exit.
The other semaphores are encapsulated in \emph{condition variables}.
\index{condition variable}%
Each condition variable also keeps track of the number of processes waiting
on the condition variable.

There are two operations on condition variables: \texttt{wait}
\index{wait}%
and
\texttt{signal}.
\index{signal}%
\autoref{fig:hoare} presents the semantics by giving an implementation of
\texttt{wait} and \texttt{signal} as well as the code that is used to enter
and exit monitor methods.
\autoref{fig:hoaretest} presents a simple example lifted directly from
the original paper~\cite{Hoare74}.

\texttt{wait} increments the condition's counter, releases the monitor mutex,
blocks while trying to procure the condition's semaphore, and after resuming
decrements the counter---in much the same way as we have seen for split binary
semaphores (SBS).
\texttt{signal} checks to see if the condition's count is non-zero, and if so
vacates the condition's semaphore---passing the baton to
a process trying to procure it---and then blocks by procuring the mutex again.
The use of \texttt{wait} and \texttt{signal} ensures that \texttt{P} and
\texttt{V} operations alternate as prescribed by the SBS technique.
Our implementations of reader/writer locks can be easily changed to
use \texttt{wait} and \texttt{signal} instead of using \texttt{P}
and \texttt{V}.
